---
title: "RSData"
author: "Oswaldo F. Domejean"
date: "October 18, 2015"
output: word_document
---

Libraries
```{r}
library(readr)
library(lubridate)
library(ggplot2)
library(dplyr)
library(vcd)
library(sqldf)
library(caret)

```

Data loading

```{r, echo=FALSE}
# Mac
setwd("~/Dropbox/Documents To Go/Kaggle/Rossman Store Sales")
#Win
setwd("C:/Users/Ing-Figueroa/Dropbox/Documents To Go/Kaggle/Rossman Store Sales")

train <- read.csv("train.csv")
test  <- read.csv("test.csv")
store <- read.csv("store.csv")
train <- merge(train,store)
test <- merge(test,store)

```

Date convertion
```{r}
fecha =as.character(train$Date)
DateConvert = strptime(fecha, format="%Y-%m-%d")
train$year = as.numeric(year(as.Date(DateConvert)))
train$month = as.numeric(month(as.Date(DateConvert)))
train$day = as.numeric(day(as.Date(DateConvert)))
# train$Date = NULL

fecha =as.character(test$Date)
DateConvert = strptime(fecha, format="%Y-%m-%d")
test$year = as.numeric(year(as.Date(DateConvert)))
test$month = as.numeric(month(as.Date(DateConvert)))
test$day = as.numeric(day(as.Date(DateConvert)))
# test$Date = NULL
```

Change number days of week by name of day of week
```{r}
train$DayOfWeek = as.factor(train$DayOfWeek)
levels(train$DayOfWeek) <- c(levels(train$DayOfWeek), "Monday")
levels(train$DayOfWeek) <- c(levels(train$DayOfWeek), "Tuesday")
levels(train$DayOfWeek) <- c(levels(train$DayOfWeek), "Wednesday")
levels(train$DayOfWeek) <- c(levels(train$DayOfWeek), "Thursday")
levels(train$DayOfWeek) <- c(levels(train$DayOfWeek), "Friday")
levels(train$DayOfWeek) <- c(levels(train$DayOfWeek), "Saturday")
levels(train$DayOfWeek) <- c(levels(train$DayOfWeek), "Sunday")
train[train$DayOfWeek == 1,2] = "Monday"
train[train$DayOfWeek == 2,2] = "Tuesday"
train[train$DayOfWeek == 3,2] = "Wednesday"
train[train$DayOfWeek == 4,2] = "Thursday"
train[train$DayOfWeek == 5,2] = "Friday"
train[train$DayOfWeek == 6,2] = "Saturday"
train[train$DayOfWeek == 7,2] = "Sunday"
train$DayOfWeek = factor(train$DayOfWeek)

test$DayOfWeek = as.factor(test$DayOfWeek)
levels(test$DayOfWeek) <- c(levels(test$DayOfWeek), "Monday")
levels(test$DayOfWeek) <- c(levels(test$DayOfWeek), "Tuesday")
levels(test$DayOfWeek) <- c(levels(test$DayOfWeek), "Wednesday")
levels(test$DayOfWeek) <- c(levels(test$DayOfWeek), "Thursday")
levels(test$DayOfWeek) <- c(levels(test$DayOfWeek), "Friday")
levels(test$DayOfWeek) <- c(levels(test$DayOfWeek), "Saturday")
levels(test$DayOfWeek) <- c(levels(test$DayOfWeek), "Sunday")
test[test$DayOfWeek == 1,3] = "Monday"
test[test$DayOfWeek == 2,3] = "Tuesday"
test[test$DayOfWeek == 3,3] = "Wednesday"
test[test$DayOfWeek == 4,3] = "Thursday"
test[test$DayOfWeek == 5,3] = "Friday"
test[test$DayOfWeek == 6,3] = "Saturday"
test[test$DayOfWeek == 7,3] = "Sunday"
test$DayOfWeek = factor(test$DayOfWeek)

train$Open = as.factor(train$Open)
test$Open = as.factor(test$Open)

train$Promo = as.factor(train$Promo)
test$Promo = as.factor(test$Promo)

train$Promo2 = as.factor(train$Promo2)
test$Promo2 = as.factor(test$Promo2)

train$SchoolHoliday = as.factor(train$SchoolHoliday)
test$SchoolHoliday = as.factor(test$SchoolHoliday)

train$year = as.factor(train$year)
test$year = as.factor(test$year)

train$month = as.factor(train$month)
test$month = as.factor(test$month)

train$day = as.factor(train$day)
test$day = as.factor(test$day)

train$CompetitionOpenSinceMonth = as.factor(train$CompetitionOpenSinceMonth)
test$CompetitionOpenSinceMonth = as.factor(test$CompetitionOpenSinceMonth)

train$CompetitionOpenSinceYear = as.factor(train$CompetitionOpenSinceYear)
test$CompetitionOpenSinceYear = as.factor(test$CompetitionOpenSinceYear)

train$Promo2SinceWeek = as.factor(train$Promo2SinceWeek)
test$Promo2SinceWeek = as.factor(test$Promo2SinceWeek)

train$Promo2SinceYear = as.factor(train$Promo2SinceYear)
test$Promo2SinceYear = as.factor(test$Promo2SinceYear)

# PromoInterval has a range of months that had been changed from ex. 'Feb,May,Aug,Nov' to 'M_2_5_8_11' and then it has been used dummy vars to create the corresponding attributes.
# For train
levels(train$PromoInterval) <- c(levels(train$PromoInterval), 'M_2_5_8_11')
train$PromoInterval[train$PromoInterval == 'Feb,May,Aug,Nov'] <- 'M_2_5_8_11'

levels(train$PromoInterval) <- c(levels(train$PromoInterval), 'M_1_4_7_10')
train$PromoInterval[train$PromoInterval == 'Jan,Apr,Jul,Oct'] <- 'M_1_4_7_10'

levels(train$PromoInterval) <- c(levels(train$PromoInterval), 'M_3_6_9_12')
train$PromoInterval[train$PromoInterval == 'Mar,Jun,Sept,Dec'] <- 'M_3_6_9_12'
train$PromoInterval <- factor(train$PromoInterval)

dmy <- dummyVars(" ~ PromoInterval", data = train)
var1 <- data.frame(predict(dmy, newdata = train))
train$PromoInterval = NULL
var1$PromoInterval. = NULL
train <- cbind(train, var1)

# For test
levels(test$PromoInterval) <- c(levels(test$PromoInterval), 'M_2_5_8_11')
test$PromoInterval[test$PromoInterval == 'Feb,May,Aug,Nov'] <- 'M_2_5_8_11'

levels(test$PromoInterval) <- c(levels(test$PromoInterval), 'M_1_4_7_10')
test$PromoInterval[test$PromoInterval == 'Jan,Apr,Jul,Oct'] <- 'M_1_4_7_10'

levels(test$PromoInterval) <- c(levels(test$PromoInterval), 'M_3_6_9_12')
test$PromoInterval[test$PromoInterval == 'Mar,Jun,Sept,Dec'] <- 'M_3_6_9_12'
test$PromoInterval <- factor(test$PromoInterval)

dmy <- dummyVars(" ~ PromoInterval", data = test)
var1 <- data.frame(predict(dmy, newdata = test))
test$PromoInterval = NULL
var1$PromoInterval. = NULL
test <- cbind(test, var1)

train$ym=as.factor(paste(train$year, train$month, sep="_"))

train$StoreAssort=as.factor(paste(train$StoreType, train$Assortment, sep="_"))
test$StoreAssort=as.factor(paste(test$StoreType, test$Assortment, sep="_"))

train$AssortStore=as.factor(paste(train$Assortment, train$StoreType, sep="_"))
test$AssortStore=as.factor(paste(test$Assortment, test$StoreType, sep="_"))

CustStoreMean = train %>%
  group_by(as.factor(Store)) %>%
  summarise(promedio=mean(Customers))

names(CustStoreMean) <- c("Store","Customers")
test <- merge(test, CustStoreMean, by = c("Store"))

#write.csv(train,"trainV1.csv",row.names=F)
#write.csv(test,"testV1.csv",row.names=F)

#train <- read.csv("trainV1.csv")
#test  <- read.csv("testV1.csv")

test$Date <- as.Date(test$Date)
train$Date <- as.Date(train$Date)

# Not sure if this is gonna be useful
train$daysinperiod <- as.integer(train$Date - min(train$Date))
test$daysinperiod <- as.integer(test$Date - min(test$Date))

# In the case of customers maybe use proportions in train and test, chek if all the stores exist in test because the proportions are going to be from customer attribute

# Fill CompetitionDistance with the average value
meanCompetitionDistance <- mean(train$CompetitionDistance, na.rm = TRUE)
train$CompetitionDistance[is.na(train$CompetitionDistance)] <- meanCompetitionDistance
meanCompetitionDistance <- mean(test$CompetitionDistance, na.rm = TRUE)
test$CompetitionDistance[is.na(test$CompetitionDistance)]   <- meanCompetitionDistance

train$CompetitionOpenSinceYear[is.na(train$CompetitionOpenSinceYear)] <- 2009
test$CompetitionOpenSinceYear[is.na(test$CompetitionOpenSinceYear)] <- 2009

train$CompetitionOpenSinceMonth[is.na(train$CompetitionOpenSinceMonth)] <- 7
test$CompetitionOpenSinceMonth[is.na(test$CompetitionOpenSinceMonth)] <- 7

train$Promo2SinceWeek[is.na(train$Promo2SinceWeek)] <- 27
test$Promo2SinceWeek[is.na(test$Promo2SinceWeek)] <- 27

train$Promo2SinceYear[is.na(train$Promo2SinceYear)] <- 2012
test$Promo2SinceYear[is.na(test$Promo2SinceYear)] <- 2012

#test[is.na(test$Open),] # These are weird
test$Open[is.na(test$Open)] <- 1  # Assume these stores are open

rm(CustStoreMean, store, var1)
rm(DateConvert, dmy, fecha)


```

Exploring the data

```{r}
str(train)
summary(train)
length(unique(train$Assortment))
# stores that are not in the test data
sum(!(unique(train$Store) %in% unique(test$Store))) 
table(train$Open) / nrow(train) # Percent Open Train
train$Open = as.factor(train$Open)
```

Exploring one variable


```{r}
unique(as.factor(train$Store))

qplot(as.factor(DayOfWeek), data=train)
qplot(Sales, data=train)
qplot(Sales, data=train, geom = "density")
qplot(Open, data = train)
qplot(Promo, data = train)
qplot(StateHoliday, data = train)
qplot(SchoolHoliday, data = train)
qplot(StoreType, data = train)
qplot(Assortment, data = train)
qplot(CompetitionDistance, data = train, geom="density")
qplot(CompetitionOpenSinceMonth, data = train)
qplot(CompetitionOpenSinceYear, data = train)
qplot(Promo2, data = train)
qplot(Promo2SinceWeek, data = train)
qplot(Promo2SinceYear, data = train)
qplot(PromoInterval, data = train)
qplot(year, data = train)
qplot(month, data = train)
qplot(day, data = train)

qplot(as.factor(DayOfWeek), data=test)
qplot(Open, data = test)
qplot(Promo, data = test)
qplot(StateHoliday, data = test)
qplot(SchoolHoliday, data = test)
qplot(StoreType, data = test)
qplot(Assortment, data = test)
qplot(CompetitionDistance, data = test)
qplot(CompetitionOpenSinceMonth, data = test)
qplot(CompetitionOpenSinceYear, data = test)
qplot(Promo2, data = test)
qplot(Promo2SinceWeek, data = test)
qplot(Promo2SinceYear, data = test)
qplot(PromoInterval, data = test)
qplot(year, data = test)
qplot(month, data = test)
qplot(day, data = test)

```

Exploring two variables
When Open is 0 (closed Store), Sales equals 0
17% in train is with Sales equals 0.
About assortment the levels a ad c have the highest sales, and maybe
divide the data set by using a and c,and other data set for b
54 stores have zero sales when they are open

```{r}

qplot(DayOfWeek, Sales, data = train)
ggplot(train, aes(x = StoreType, y = Sales)) + geom_bar(stat = "identity")
ggplot(train, aes(x = Assortment, y = Sales)) + geom_bar(stat = "identity")
qplot(Promo, Sales, data = train)
qplot(Promo2, Sales, data = train)
qplot(Promo, Sales/10, data = train)
qplot(Promo2, Sales/10, data = train)
qplot(StateHoliday, Sales/10, data = train)
qplot(SchoolHoliday, Sales/10, data = train)

# Competition distance vs sales
qplot(CompetitionDistance/100, Sales/100, data = train, geom = c("point", "smooth"))

```

Exploring three variables
```{r}

structable(StoreType+Assortment~Sales, train)
table(train$Open, train$StateHoliday)
table(train$StateHoliday, train$SchoolHoliday, train$Open)
structable(SchoolHoliday+StateHoliday~Open, train)

train %>%
  group_by(DayOfWeek) %>%
  summarise(total=sum(Sales/100))

train %>%
  group_by(Assortment) %>%
  summarise(total=sum(Sales/10))

train %>%
  group_by(StoreType) %>%
  summarise(total=sum(Sales/10))

train %>%
  group_by(Assortment) %>%
  summarise(total=mean(Sales))

train %>%
  group_by(StoreType) %>%
  summarise(total=mean(Sales))

select(train, Date, StoreType, Sales) %>%
  filter(Sales != 0) %>%
  group_by(as.Date(Date), StoreType) %>%
  summarise(total=sum(Sales/10))

train %>%
  group_by(Open) %>%
  summarise(total=mean(Sales))

train %>%
  group_by(Promo2) %>%
  summarise(total=sum(Sales/10))

select(train, StoreType, Promo, Sales) %>%
  group_by(StoreType, Promo) %>%
  summarise(total=sum(Sales))

select(train, Assortment, Promo, Sales) %>%
  group_by(Assortment, Promo) %>%
  summarise(total=sum(Sales))

select(train, StoreType, Assortment, Sales) %>%
  group_by(StoreType, Assortment) %>%
  summarise(total=sum(Sales))

select(train, Assortment, StoreType, Sales) %>%
  group_by(Assortment, StoreType) %>%
  summarise(total=sum(Sales))

select(train, StateHoliday, StoreType, Sales) %>%
  group_by(StateHoliday, StoreType) %>%
  summarise(total=sum(Sales))

select(train, StoreType, StateHoliday, Sales) %>%
  group_by(StoreType, StateHoliday) %>%
  summarise(total=sum(Sales))

train %>%
  filter(Sales != 0) %>%
  ggplot(aes(x=as.Date(Date), y = Sales, color = StoreType)) +
  geom_smooth(size = 2)

train %>%
  filter(Customers != 0) %>%
  ggplot(aes(x=as.Date(Date), y = Customers, color = StoreType)) +
  geom_smooth(size =2)

train %>%
  filter(Sales != 0) %>%
  ggplot(aes(x = as.Date(Date), y = Sales, color = Assortment)) +
  geom_smooth(size = 2)

train %>%
  filter(Customers != 0) %>%
  ggplot(aes(x = as.Date(Date), y = Customers, color = Assortment)) +
  geom_smooth(size = 2)

alfa = select(train, year, month, Sales) %>%
  group_by(month, year) %>%
  summarise(total = sum(Sales))

ggplot(data=alfa,
       aes(x=month, y=total, color=year)) +
       geom_line(aes(group=year))

train %>%
  filter(Sales != 0) %>%
  ggplot(aes(x = as.Date(Date), y = Sales, color = Assortment)) +
  geom_smooth(size = 2)

# Total sales by StoreType and Assortment
select(train, ym, StoreType, Sales) %>%
  filter(Sales != 0) %>%
  group_by(ym, StoreType) %>%
  summarise(total = sum(Sales)) %>%
  ggplot(aes(x = ym, y = total, group = StoreType, color = StoreType)) +
  geom_line() + geom_point()

select(train, ym, Assortment, Sales) %>%
  filter(Sales != 0) %>%
  group_by(ym, Assortment) %>%
  summarise(total = sum(Sales)) %>%
  ggplot(aes(x = ym, y = total, group = Assortment, color = Assortment)) +
  geom_line() + geom_point()

# Mean sales by StoreType and Assortment
select(train, ym, StoreType, Sales) %>%
  filter(Sales != 0) %>%
  group_by(ym, StoreType) %>%
  summarise(total = mean(Sales)) %>%
  ggplot(aes(x = ym, y = total, group = StoreType, color = StoreType)) +
  geom_line() + geom_point()

select(train, ym, Assortment, Sales) %>%
  filter(Sales != 0) %>%
  group_by(ym, Assortment) %>%
  summarise(total = mean(Sales)) %>%
  ggplot(aes(x = ym, y = total, group = Assortment, color = Assortment)) +
  geom_line() + geom_point()

# Total sales by month, year StoreType and Assortment
select(train, year, month, StoreType, Sales) %>%
  filter(Sales != 0) %>%
  group_by(month, year, StoreType) %>%
  summarise(suma = sum(Sales)) %>%
  ggplot(aes(x=month, y=suma, color=StoreType)) +
  geom_line(aes(group=StoreType))

select(train, year, month, Assortment, Sales) %>%
  filter(Sales != 0) %>%
  group_by(month, year, Assortment) %>%
  summarise(suma = sum(Sales)) %>%
  ggplot(aes(x=month, y=suma, color=Assortment)) +
  geom_line(aes(group=Assortment))

# Total sum by StoreAssort that is a combination between StoreType and Assortment
select(train, ym, StoreAssort, Sales) %>%
  filter(Sales != 0) %>%
  group_by(ym, StoreAssort) %>%
  summarise(total = sum(Sales)) %>%
  ggplot(aes(x = ym, y = total, group = StoreAssort, color = StoreAssort)) +
  geom_line() + geom_point()

# Total sales by month, year with StoreAssort
select(train, year, month, StoreAssort, Sales) %>%
  filter(Sales != 0) %>%
  group_by(month, year, StoreAssort) %>%
  summarise(suma = sum(Sales)) %>%
  ggplot(aes(x = month, y = suma, color = StoreAssort)) +
  geom_line(aes(group=StoreAssort, linetype=StoreAssort))

# Total sales by day with StoreAssort
select(train, day, StoreAssort, Sales) %>%
  filter(Sales != 0) %>%
  group_by(day, StoreAssort) %>%
  summarise(suma = sum(Sales)) %>%
  ggplot(aes(x = day, y = suma, color = StoreAssort)) +
  geom_line(aes(group=StoreAssort, linetype=StoreAssort))

# Total sum by AssortStore that is a combination between StoreType and Assortment
select(train, ym, AssortStore, Sales) %>%
  filter(Sales != 0) %>%
  group_by(ym, AssortStore) %>%
  summarise(total = sum(Sales)) %>%
  ggplot(aes(x = ym, y = total, group = AssortStore, color = AssortStore)) +
  geom_line() + geom_point()

sqldf(
  'select count(*)
  from train
  where Open = "1"
  and Sales = 0')

sqldf('select Assortment, StoreType, sum(Sales)/10
      from train
      group by Assortment, StoreType')

sqldf('select Promo, sum(Sales)/10
      from train
      group by Promo')



```


Building the datasets

```{r}

train$Date = NULL
test$Date = NULL
train$StoreType = NULL
test$StoreType = NULL
train$Assortment = NULL
test$Assortment = NULL
train$ym = NULL


test$ym = NULL

trainG1 = train %>%
  filter(Open == 1 & Sales != 0 & StoreAssort == "a_a")
trainG1$Open = NULL
trainG1$StoreAssort = NULL
write.csv(trainG1, "trainG1.csv", row.names = FALSE)

testG1 = test %>%
  filter(Open == 1 & StoreAssort == "a_a")
testG1$Open = NULL
testG1$StoreAssort = NULL
write.csv(testG1, "testG1.csv", row.names = FALSE)

```

